---
title: "Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(naniar)
library(shiny)
library(shinydashboard)
library(shinythemes)
```

```{r}
srate <- readr::read_csv("master.csv")
```


```{r}
srate %>%
  summarise(totcountry = n_distinct(country))
```
```{r}
srate %>%
  arrange(desc(year))
```



```{r}
sratee <- srate %>%
  rename(gdpyr = `gdp_for_year ($)`, gdpcap = `gdp_per_capita ($)`, suicper100k = `suicides/100k pop`)
```

```{r}
sratee$country <- as.factor(sratee$country)
sratee$generation <- as.factor(sratee$generation)
sratee$age <- as.factor(sratee$age)
str(sratee)
```

```{r}
sratee %>%
  select(country, year, suicper100k) %>%
  pivot_wider(
    names_from = country,
    values_from = suicper100k,
    values_fn = list(suicper100k = sum))
```
```{r}
socialist <- sratee %>%
  select(country, year, suicper100k) %>%
  filter(country == "Finland" | country == "Denmark" | country == "Netherlands" | country == "Canada" | country == "Sweden" | country == "Norway" | country == "Ireland" | country == "New Zealand" | country == "Belgium") %>%
   pivot_wider(
    names_from = year, 
    values_from = suicper100k,
    values_fn = list(suicper100k = sum)) %>%
  pivot_longer(
    `1985`:`2016`,
    names_to = "year",
    values_to = "count"
  )
```

```{r}
sratee %>%
  select(country, year, suicper100k) %>%
  filter(country == "Finland" | country == "Denmark" | country == "Netherlands" | country == "Canada" | country == "Sweden" | country == "Norway" | country == "Ireland" | country == "New Zealand" | country == "Belgium", year == "2013") %>%
  pivot_wider(
    names_from = year,
    values_from = suicper100k,
    values_fn = list(suicper100k = sum)) %>%
  ggplot(aes(x = country, y = `2013`)) +
  geom_bar(stat = "identity")
```

```{r}
sratee %>%
  select(country, year, suicper100k) %>%
  filter(country == "Singapore" | country == "Switzerland" | country == "Australia" | country == "Canada" | country == "Ireland" | country == "United Kingdom" | country == "United Arab Emirates" | country == "New Zealand" | country == "Taiwan", year == "2013") %>%
  pivot_wider(
    names_from = year,
    values_from = suicper100k,
    values_fn = list(suicper100k = sum)) %>%
  ggplot(aes(x = country, y = `2013`)) +
  geom_bar(stat = "identity")
```

```{r}
sratesex <- sratee %>%
  select(sex, suicides_no, population) %>%
  pivot_wider(
    names_from = sex, 
    values_from = suicides_no,
    values_fn = list(suicides_no = sum)) 
sratesex
```

```{r}
sratesex %>%
  summarise(totsuicmale = sum(male), totsuicf = sum(female))
```

```{r}
srateegen <- sratee %>%
  select(generation, suicides_no)
```

Suicides by generation, is skewed by the number of people within the generations
```{r}
srateegen %>%
  ggplot(aes(x = generation, y = suicides_no, fill = generation))+
  geom_bar(stat = "identity")
```
1901 to 1927 GI
1928 to 1945 Silent
1946 to 1964 Boomers
1965 to 1980 Gen X
1981 to 1996 Millenials
no consensus for beginning 2012 

DO POPULATION BY GENERATION



```{r}
srateeage <- sratee %>%
  select(suicides_no, age, population)
```


Suicides by age group
```{r warning = FALSE}
srateeage %>%
  ggplot(aes(x= age, y = suicides_no, fill = age))+
  geom_bar(stat = "identity")
```



Total number of sampled people
```{r}
srateeage %>%
  ggplot(aes(x = age, y = population, fill = age))+
  geom_bar(stat = "identity")
```
```{r}
srateeage %>%
  group_by(age) %>%
  pivot_wider(
    names_from = age,
    values_from = population,
    values_fn = list(population = sum))
```


```{r}
sratee %>%
  select(generation, suicides_no, year, age) %>%
  filter(generation == "Generation Z") %>%
  arrange(age, year)
```

App for showing suicides by generation and suicides by gender
```{r}
ui <- 
  dashboardPage(
  dashboardHeader(title = "Suicide-Generation App"),
  dashboardSidebar(disable = TRUE),
  dashboardBody(
  fluidRow(
  box(title = "Plot Options", background = "maroon", solidHeader = TRUE, width = 2,
  selectInput("generation", "Generation", choices = c("G.I. Generation", "Silent", "Boomers", "Generation X", "Millenials", "Generation Z"),
              selected = "G.I. Generation"),
  selectInput("year", "Year", choices = c("1985", "1990", "1995", "2000", "2005", "2010", "2016"),
              selected = "1985"),
  selectInput("sex", "Sex", choices = c("male", "female"))),
  box(title = "Suicide Generation Plot", background = "maroon", solidHeader = TRUE, width = 10,
  plotOutput("plot1", width = "1657px", height = "400px"))),
  fluidRow(
  box(title = "Plot Options", background = "fuchsia", solidHeader = TRUE, width = 2, 
  selectInput("yearr", "Year", choices = c("1985", "1990", "1995", "2000", "2005", "2010", "2016"),
              selected = "1985"),
  selectInput("agee", "Age", choices = c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"), selected = "5-14 Years")),
  box(title = "Suicide Sex Plot", background = "fuchsia", solidHeader = TRUE, width = 10,
    plotOutput("plot2", width = "1657px", height = "400px"
)
)
)
)
)

server <- function(input, output, session) { 
  
  # the code to make the plot of iris data grouped by species
  output$plot1 <- renderPlot({
    
    sratee %>%
      filter(sex == input$sex, year == input$year, generation == input$generation) %>%
      ggplot(aes(x = age, y = suicides_no, fill = age))+
      geom_bar(stat = "identity")
  })
  output$plot2 <- renderPlot({
  
    sratee %>%
      filter(year == input$yearr, age == input$agee) %>%
      ggplot(aes(x = sex, y = suicides_no, fill = sex))+
      geom_bar(stat = "identity")
  })
  # stop the app when we close it
  session$onSessionEnded(stopApp)

  }

shinyApp(ui, server)
```
Do you want the year to be controlled by one button in order to show sex differences in accordance with generation?, or have them as completely separate graphs?



Year on the x axis, number of suicides on the y axis

```{r}
srateeyear <- sratee %>%
  select(suicides_no, year) %>%
  filter(year != "2016") %>%
  group_by(year) %>%
  summarise(suicides = sum(suicides_no)) 
```

I omitted the year 2016 from the data set for this since it appears it wasn't sampeled throughout the year. It was heavily skewing my data.
Suicides by the year
```{r}
srateeyear %>%
  ggplot(aes(x = year, y = suicides))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```

GDPpercap vs suicide 
```{r}
sratee %>%
  select(gdpcap, suicides_no) %>%
  group_by(gdpcap) %>%
  summarise(suicides = sum(suicides_no)) %>%
  arrange(gdpcap) %>%
  ggplot(aes(x = gdpcap, y = suicides)) +
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```



```{r}
suicide %>%
  select(`gdp_per_capita_($)`, country, suicide_number) %>%
  group_by(country) %>%
  summarise(meangdp = mean(`gdp_per_capita_($)`)) %>%
  top_n(-10) %>%
  arrange(desc(meangdp))
```

```{r}
suicide %>%
  select(`gdp_per_capita_($)`, country, suicide_number) %>%
  group_by(country) %>%
  summarise(meangdp = mean(`gdp_per_capita_($)`)) %>%
  top_n(10) %>%
  arrange(desc(meangdp))
```



```{r}
suicidetop10 <- suicide %>%
  select(country, suicide_number, population) %>%
  filter(country == "United States" | country == "Luxembourg" | country == "Qatar" | country == "Switzerland" | country == "Norway" | country == "San Marino" | country == "Denmark" | country == "United Arab Emirates" | country == "Sweden" | country == "Iceland") %>%
  group_by(country) %>%
  summarise(totsuic = sum(suicide_number), totpop = sum(population)) %>%
  mutate(suiciderate = totsuic/totpop)
```

```{r}
suicidetop10 %>%
  select(suiciderate)
```



```{r}
suicidebot10 <- suicide %>%
  select(country, suicide_number, population) %>%
  filter(country == "Guyana" | country == "Ukraine" | country == "Albania" | country == "Philippines" | country == "Azerbaijan" | country == "Uzbekistan" | country == "Sri Lanka" | country == "Kyrgyzstan" | country == "Dominica" | country == "Kiribati") %>%
  group_by(country) %>%
  summarise(totsuic = sum(suicide_number), totpop = sum(population)) %>%
  mutate(suiciderate = totsuic/totpop)
```

```{r}
suicidebot10 %>%
  select(suiciderate)
```

```{r}
suicideratedf<-data.frame("Group" = c("Lower"), "Suicide Rate" = c(3.160827e-05, 1.481345e-05, 0.000000e+00, 2.064528e-04, 7.146652e-05, 1.136539e-04, 2.002689e-05, 3.048394e-04, 2.487040e-04, 7.154891e-05))
```
```{r}
top10 <- data.frame("Group" = "Upper", "Suicide Rate" = c(1.363585e-04, 1.317808e-04, 1.511601e-04, 1.327779e-04, 2.624535e-05, 5.074532e-05, 1.492121e-04, 1.747870e-04, 1.704003e-05, 1.283846e-04))
```

```{r}
comparetest <- rbind(suicideratedf, top10)
```

```{r}
t_test <- t.test(Suicide.Rate~Group, data = comparetest, alternative = "less", conf.level = 0.95)
print(t_test)
```







