---
title: "Project: Suicide in Global Scale"
author: "Lily Chen"
date: "2/28/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
#install.packages(jpeg)
#intsall.packages(naniar)
#intsall.packages(skimr)
#intsall.packages(tidyverse)
#intsall.packages(shiny)
#intsall.packages(shinydashboard)
```


## Introduction


## Reference
-United Nations Development Program. (2018). Human development index (HDI). Retrieved from http://hdr.undp.org/en/indicators/137506

-World Bank. (2018). World development indicators: GDP (current US$) by country:1985 to 2016. Retrieved from http://databank.worldbank.org/data/source/world-development-indicators#[Szamil]. (2017). 

-Suicide in the Twenty-First Century [dataset]. Retrieved from https://www.kaggle.com/szamil/suicide-in-the-twenty-first-century/notebook

-World Health Organization. (2018). Suicide prevention. Retrieved from http://www.who.int/mental_health/suicide-prevention/en/


## Library
Here are the libraries that we use for analyzing the data.
```{r echo=TRUE, message=FALSE}
library(jpeg)
library(naniar)
library(skimr)
library(tidyverse)
library(shiny)
library(shinydashboard)
```

## Load Data
```{r echo=TRUE, message=FALSE}
suicide <- readr::read_csv("data/master.csv")
```

## Data Overview
```{r echo=TRUE}
glimpse(suicide)
```

## NAs Overview
```{r echo=FALSE}
naniar::miss_var_summary(suicide)
```

## Clean Dara
```{r echo=TRUE, message=FALSE}
names(suicide) <- str_replace_all(names(suicide), c(" " = "_", "/" = "_", "-" = "_"))
```

```{r echo=TRUE, message=FALSE}
suicide <-
suicide %>% 
  dplyr::rename(
    suicide_number = suicides_no,
    suicide_per_100kpop = suicides_100k_pop)
```

```{r echo=TRUE, message=FALSE}
suicide$country <- as.factor(suicide$country)
suicide$age <- as.factor(suicide$age)
suicide$sex <- as.factor(suicide$sex)
suicide$country_year <- as.factor(suicide$country_year)
suicide$generation <- as.factor(suicide$generation)
suicide$year <- as.factor(suicide$year)
```

## Data Overview
```{r echo=TRUE}
glimpse(suicide)
```

## Group by Year: Suicide Number and Suicide Rate 
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population))
```

## Group by Year: Suicide Number
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population)) %>% 
  ggplot(aes(x = year, y = suicide_number)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Year vs. Suicide Number",
       y = "suicide number")
```

## Group by Year: Population
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population)) %>% 
  ggplot(aes(x = year, y = population)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Year vs. Population",
       y = "population")
```

## Group by Year: Suicide Rate
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population)) %>% 
  ggplot(aes(x = year, y = suicide_rate)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Year vs. Suicide Rate",
       y = "suicide rater")
```

## Group by Sex: Suicide Number and Suicide Rate
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(sex) %>% 
  summarize(num_suicide = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (num_suicide)/(population))
```

## Group by Sex: Suicide Number
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(sex) %>% 
  summarize(num_suicide = sum(suicide_number)) %>% 
  ggplot(aes(x = sex, y = num_suicide)) +
  geom_bar (stat = "identity") +
  labs(title = "Sex vs. Suicide Number",
       y = "suicide number")
```

## Group by Sex: Population
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(sex) %>% 
  summarize(population = sum(population)) %>% 
  ggplot(aes(x = sex, y = population)) +
  geom_bar (stat = "identity") +
  labs(title = "Sex vs. Population",
       y = "population")
```

## Group by Sex: Suicide Rate
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(sex) %>% 
  summarize(num_suicide = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (num_suicide)/(population)) %>%  
  ggplot(aes(x = sex, y = suicide_rate)) +
  geom_bar (stat = "identity") +
  labs(title = "Sex vs. Suicide Rate",
       y = "suicide rate")
```


## Group by Age: Suicide Number and Suicide Rate
```{r echo=FALSE, message=FALSE}
suicide %>% 
  mutate(age_class=case_when(age=="5-14 years" ~ "A(5-14)", age=="15-24 years" ~ "B(15-24)", age=="25-34 years" ~ "C(25-34)", age=="35-54 years" ~ "D(35-45)", age=="55-74 years" ~ "E(55-74)",age=="75+ years" ~ "F(75+)" )) %>% 
  group_by(age_class) %>% 
  summarize(num_suicide = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (num_suicide)/(population))
```

## Group by Age: Suicide Number
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(age) %>% 
  summarize(num_suicide = sum(suicide_number)) %>% 
  ggplot(aes(x = age, y = num_suicide)) +
  geom_bar (stat = "identity") +
  scale_x_discrete(limits=c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
  labs(title = "Age vs. Suicide Number",
       y = "suicide number")
```

## Group by Age: Population
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(age) %>% 
  summarize(population = sum(population)) %>% 
  ggplot(aes(x = age, y = population)) +
  geom_bar (stat = "identity") +
  scale_x_discrete(limits=c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
  labs(title = "Age vs. Population",
       y = "population")
```

## Group by Age: Suicide Rate
```{r echo=FALSE, message=FALSE}
suicide %>% 
  group_by(age) %>% 
  summarize(num_suicide = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (num_suicide)/(population)) %>% 
  ggplot(aes(x = age, y = suicide_rate)) +
  geom_bar (stat = "identity") +
  scale_x_discrete(limits=c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
  labs(title = "Age vs. Suicide Rate",
       y = "suicide rate")
```


## Suicide Rate vs. GDP Per Capita
```{r echo=FALSE, message=FALSE}
suicide %>% 
  ggplot(aes(x = `gdp_per_capita_($)`, y = suicide_per_100kpop)) +
  geom_jitter () +
  labs(title = "GDP Per Capita vs. Suicide Per 100k Popullation",
       x = "GDP Per Capita",
       y = "Suicide Per 100k Population")
```

##Shiny App 1
```{r include=FALSE}
ui <- dashboardPage(
  dashboardHeader(title = "Suicide Per 100k Pop by Age Group"),
  dashboardSidebar(disable = T),
  dashboardBody(
    fluidRow(
      box(title = "Plot Options", width = 3,
          selectInput("x", "Select Year", choices = unique(suicide$year), 
                       selected = "2010"),
          selectInput("y", "Select Country", choices = unique(suicide$country),
                      selected = "Albania"),
          selectInput("z", "Select Sex", choices = unique(suicide$sex),
                      selected = "Male")
      ), # close the first box
      box(title = "Suicide Per 100k Pop by Age Group", width = 7,
          plotOutput("plot", width = "600px", height = "500px")
      ) # close the second box
    ) # close the row
  ) # close the dashboard body
) # close the ui

server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
    suicide %>% 
      filter(year==input$x & country==input$y & sex==input$z) %>% 
      ggplot(aes(x=age, y=suicide_per_100kpop)) + 
      geom_col(color="black", fill="steelblue", alpha=0.75) +
      theme_light(base_size = 18) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(x = "Age Group", y = "Suicide Per 100k Pop")
  })
  
  # stop the app when we close it
  session$onSessionEnded(stopApp)
  
}

shinyApp(ui, server)
```



