---
title: "Project"
author: "Lily Chen"
date: "2/28/2020"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


**1. Load data and library. (tidyverse, naniar, skimr, visdat)**

```{r}
library(naniar)
library(skimr)
library(tidyverse)
library(shiny)
library(shinydashboard)
library(jpeg)
```

```{r}
getwd()
```


```{r}
suicide <- readr::read_csv("data/master.csv")
```

**2. Use the function(s) of your choice to get an idea of the overall structure of the data frame, including its dimensions, column names, variable classes, etc. As part of this, determine how NAs are treated in the data. Summarize the number and percentage of NAs in each column. Visualize NAs.**

```{r}
glimpse(suicide)
```

```{r}
suicide %>% 
  summarize(number_nas = sum(is.na(suicide)))
```


```{r}
naniar::miss_var_summary(suicide)
```

**3. Make necessary changes to clean the data. (Tidy data, change classes of variable, rename column names, replace NAs, etc.)**

```{r}
names(suicide) <- str_replace_all(names(suicide), c(" " = "_", "/" = "_", "-" = "_"))
```

```{r}
suicide <-
suicide %>% 
  dplyr::rename(
    suicide_number = suicides_no,
    suicide_per_100kpop = suicides_100k_pop)
```

```{r}
suicide$country <- as.factor(suicide$country)
suicide$age <- as.factor(suicide$age)
suicide$sex <- as.factor(suicide$sex)
suicide$country_year <- as.factor(suicide$country_year)
suicide$generation <- as.factor(suicide$generation)
suicide$year <- as.factor(suicide$year)
```


```{r}
glimpse(suicide)
```

**4. How has global suicide number, population, suicide rate changed between 1985 and 2016?**

```{r}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population))
```


```{r}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population)) %>% 
  ggplot(aes(x = year, y = suicide_number)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Year vs. Suicide Number",
       y = "suicide number")

```

```{r}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population)) %>% 
  ggplot(aes(x = year, y = population)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Year vs. Population",
       y = "population")
```

```{r}
suicide %>% 
  group_by(year) %>% 
  summarize(suicide_number = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (suicide_number)/(population)) %>% 
  ggplot(aes(x = year, y = suicide_rate)) +
  geom_point() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(title = "Year vs. Suicide Rate",
       y = "suicide rater")
```

**5. Summarize suicide number and suicide rate bewteen female and male.**

```{r}
suicide %>% 
  group_by(sex) %>% 
  summarize(num_suicide = sum(suicide_number), population = sum(population), suicide_per_100kpop = sum(suicide_per_100kpop), suicide_rate = (num_suicide)/(population))
```

```{r}
suicide %>% 
  group_by(sex) %>% 
  summarize(num_suicide = sum(suicide_number)) %>% 
  ggplot(aes(x = sex, y = num_suicide)) +
  geom_bar (stat = "identity") +
  labs(title = "Sex vs. Suicide Number",
       y = "suicide number")
```


```{r}
suicide %>% 
  group_by(sex) %>% 
  summarize(population = sum(population)) %>% 
  ggplot(aes(x = sex, y = population)) +
  geom_bar (stat = "identity") +
  labs(title = "Sex vs. Population",
       y = "population")
```


**6. Summarize suicide number and suicide rate among different age groups.**

```{r}
suicide %>% 
  mutate(age_class=case_when(age=="5-14 years" ~ "A(5-14)", age=="15-24 years" ~ "B(15-24)", age=="25-34 years" ~ "C(25-34)", age=="35-54 years" ~ "D(35-45)", age=="55-74 years" ~ "E(55-74)",age=="75+ years" ~ "F(75+)" )) %>% 
  group_by(age_class) %>% 
  summarize(num_suicide = sum(suicide_number))
```


```{r}
suicide %>% 
  group_by(age) %>% 
  summarize(num_suicide = sum(suicide_number)) %>% 
  ggplot(aes(x = age, y = num_suicide)) +
  geom_bar (stat = "identity") +
  scale_x_discrete(limits=c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
  labs(title = "Age vs. Suicide Number",
       y = "suicide number")
```

```{r}
suicide %>% 
  group_by(age) %>% 
  summarize(population = sum(population)) %>% 
  ggplot(aes(x = age, y = population)) +
  geom_bar (stat = "identity") +
  scale_x_discrete(limits=c("5-14 years", "15-24 years", "25-34 years", "35-54 years", "55-74 years", "75+ years"))
  labs(title = "Age vs. Population",
       y = "population")
```


**8. Summarize relationship between suicide rate and gdp per capita**

```{r}
suicide %>% 
  ggplot(aes(x = `gdp_per_capita_($)`, y = suicide_per_100kpop)) +
  geom_jitter () +
  labs(title = "GDP Per Capita vs. Suicide Per 100k Popullation",
       x = "GDP Per Capita",
       y = "Suicide Per 100k Population")
```

**9. Shiny App**
```{r}
ui <- dashboardPage(
  dashboardHeader(title = "Suicide Per 100k Pop by Age Group"),
  dashboardSidebar(disable = T),
  dashboardBody(
    fluidRow(
      box(title = "Plot Options", width = 3,
          selectInput("x", "Select Year", choices = unique(suicide$year), 
                       selected = "2010"),
          selectInput("y", "Select Country", choices = unique(suicide$country),
                      selected = "Albania"),
          selectInput("z", "Select Sex", choices = unique(suicide$sex),
                      selected = "Male")
      ), # close the first box
      box(title = "Suicide Per 100k Pop by Age Group", width = 7,
          plotOutput("plot", width = "600px", height = "500px")
      ) # close the second box
    ) # close the row
  ) # close the dashboard body
) # close the ui

server <- function(input, output, session) { 
  
  output$plot <- renderPlot({
    suicide %>% 
      filter(year==input$x & country==input$y & sex==input$z) %>% 
      ggplot(aes(x=age, y=suicide_per_100kpop)) + 
      geom_col(color="black", fill="steelblue", alpha=0.75) +
      theme_light(base_size = 18) +
      theme(axis.text.x = element_text(angle = 60, hjust = 1))+
      labs(x = "Age Group", y = "Suicide Per 100k Pop")
  })
  
  # stop the app when we close it
  session$onSessionEnded(stopApp)
  
}

shinyApp(ui, server)
```



